Higher-Order unification for free!

### 2024-04-12 paper/main.tex
git url update

@@

\paragraph{Contributions}
In this paper we identify a minimal language \Fo{} in which the problems
sketched in the introduction can be formally described.
We detail an encoding of a logic program in \Fo{} to a strongly related
logic program in \Ho (the language of the meta-language) and we show that
the higher-order unification procedure of the meta language \Ue{} can be
efficiently used to simulate a higher-order unification procedure \Uo for
the object language that features $\eta\beta$-conversion. We show how \Uo
can be extended with heuristics to deal with problems outside the pattern
fragment.\\
Section~\ref{sec:problem-statement} formally states the problem and gives the
intuition behind our solution; \cref{sec:simulation} sets up a basic
simulation of first-order logic programs, \cref{sec:llam} and \cref{sec:eta}
extend it to higher-order logic programs in the pattern fragment
while \cref{sec:beta} goes beyond the pattern fragment.
Section \ref{sec:implementation} discusses the implementation in Elpi.\\
The $\lambda$Prolog code discussed in the paper can be accessed at the
address \url{https://github.com/FissoreD/paper-ho}.

>>>>

\def\githubUrl{\url{https://github.com/FissoreD/ho-unif-for-free}}

@@

\paragraph{Contributions}
In this paper we identify a minimal language \Fo{} in which the problems
sketched in the introduction can be formally described.
We detail an encoding of a logic program in \Fo{} to a strongly related
logic program in \Ho (the language of the meta-language) and we show that
the higher-order unification procedure of the meta language \Ue{} can be
efficiently used to simulate a higher-order unification procedure \Uo for
the object language that features $\eta\beta$-conversion. We show how \Uo
can be extended with heuristics to deal with problems outside the pattern
fragment.\\
Section~\ref{sec:problem-statement} formally states the problem and gives the
intuition behind our solution; \cref{sec:simulation} sets up a basic
simulation of first-order logic programs, \cref{sec:llam} and \cref{sec:eta}
extend it to higher-order logic programs in the pattern fragment
while \cref{sec:beta} goes beyond the pattern fragment.
Section \ref{sec:implementation} discusses the implementation in Elpi.\\
The $\lambda$Prolog code discussed in the paper can be accessed at the
address \githubUrl.
